timesliced_random_cell_selection

### 2018-12-20 src/timesliced_random_cell_selection.Rmd
background update

```{r}
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand20k/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 20000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

@@

Second you need to put the background (background_all) into a spatialpointsdataframe.
```{r}
xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)
```

@@

    }
}
write.csv(background_all_sp, "../output/bio/background_complete.csv", row.names = FALSE)
background_all <- as.data.frame(background_all_sp)
head(background_all_sp)
```

@@

```{r}
backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
backna
background_all[complete.cases(background_all), ] #removes NA values
write.csv(background_all, "../output/bio/background_complete_na.csv", row.names = FALSE)
```

randomly select 10000 rows from each timeslice
```{r}
#background_all <- read.csv("../output/bio/background_complete_na.csv", header = TRUE)
#split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)

@@

```{r}
bckrnd_folder <- "../output/bio/background/rand20k/no_na/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- ". ./output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample 
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

# observation occurrence yes or no

create a new column for Occurence, and populate with "0" (because these are all background)
```{r}
background_all <- read.csv("../output/bio/background_complete.csv")
background_all$occurrence <- "0"
write.csv(background_all, "../output/bio/background_complete_obs.csv", row.names = FALSE)
head(background_all)
```

>>>>

```{r}
bckrnd_folder <- "../output/bio/background/raw/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand20k/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 10000) #where 20000 = number of rows to sample (more than 10k as some points will have NA values due to depth - I will select 10000 of these without NA values to create a large sample as per maxent)
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

@@

Second you need to put the background (background_all) into a spatialpointsdataframe.
```{r}
background_all2 <- background_all[background_all$year == 2007, ]
xy <- background_all2[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude

#xy <- background_all[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
background_all_sp <- SpatialPointsDataFrame(coords = xy, data = background_all2, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
head(background_all_sp)
```

@@

    }
}
write.csv(background_all_sp, "../output/bio/background_complete_2007.csv", row.names = FALSE)
background_all <- as.data.frame(background_all_sp)
head(background_all_sp)
```

@@

```{r}
backna <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
backna
background_all <- background_all[complete.cases(background_all), ] #removes NA values
backna2 <- sum(!complete.cases(background_all)) #this just counts the number of rows that contain NA values
backna2
write.csv(background_all, "../output/bio/background_complete_na.csv", row.names = FALSE)
```

randomly select 10000 rows from each timeslice
```{r}
#background_all <- read.csv("../output/bio/background_complete_na.csv", header = TRUE)
split_cell_back <- split(background_all, list(background_all$year, background_all$month), drop = TRUE)

@@

```{r}
bckrnd_folder <- "../output/bio/background/rand20k/no_na/"      # path to folder that holds multiple .csv files
bckrnd_list <- list.files(path = bckrnd_folder, pattern="*.csv") # create list of all .csv files in folder
randbckoutput <- "../output/bio/background/rand20k/no_na_10krand/" #where the files are to be saved to
# read in each .csv file in file_list and create a data frame with the same name as the .csv file
for (i in 1:length(bckrnd_list)){
  sel <- read.csv(paste(bckrnd_folder, bckrnd_list[i], sep=''))
  sel <- sample_n(sel, 10000) #where 10000 = number of rows to sample 
  write.csv(sel, file = paste0(randbckoutput, bckrnd_list[i]), row.names = FALSE) #dont need to add .csv as it's already in the filename
}
head(sel)
```

crap - the years from 2007 onwards have less than 10k points to choose from...
create a random 30k for each of these month/year periods

```{r}
filenames  <- list.files(path = randbckoutput, pattern="*.csv", full.names = TRUE) 
background_all <- rbindlist(lapply(filenames,fread, fill = TRUE))
write.csv(background_all, "../output/bio/background_all_complete_na2.csv", row.names = FALSE)
head(background_all)
```
# observation occurrence yes or no

create a new column for Occurence, and populate with "0" (because these are all background)
```{r}
background_all$occurrence <- "0"
write.csv(background_all, "../output/bio/background_complete_obs.csv", row.names = FALSE)
head(background_all)
```
